
---
title: "Using ClearScatterplot"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using ClearScatterplot}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(MultiModalGraphics)
library(dplyr)
```

## Introduction

This vignette demonstrates how to use the `ClearScatterplot` class to create enhanced scatterplots. The `ClearScatterplot` class provides functionalities for creating scatterplots with enhanced visualization features like coloring based on significance levels and plotting against different variables.

## Example: Creating a Scatterplot

First, load the necessary data and create an instance of `ClearScatterplot`.

```{r load-data}
plotdata <- get_clear_scatterplot_df()

scatterplotObject <- ClearScatterplot(data = plotdata,
                        logFoldChange = "log2fc",
                        timePointColumn = "timePoint",
                        timePointLevels = c("T10R1", "T5R1")
                        )
```

Next, create the scatterplot using the `createPlot` method, specifying various parameters such as color and thresholds.

```{r create-plot}
scattered_plot <- createPlot(scatterplotObject, color1 = "cornflowerblue", color2 = "grey", color3="indianred",
                             highLog2fc = 0.585, lowLog2fc = -0.585, negLog10pValue = 1.301,
                             expressionDirection = "regulation", negativeLogPValue="negLog10p",
                             timeVariable="reg_time_org", xAxis = "organ", yAxis = "timePoint")
```

Finally, display the plot. This will call the 'show' method to render the plot.

```{r display-plot}
scattered_plot  # Display the plot
```

## Session info
```{r, echo=FALSE}
sessionInfo()
