---
title: "ThresholdedScatterplot Example"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ThresholdedScatterplot Example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Abstract

`ThresholdedScatterplot` is a highly customizable scatter plot class that extends `ggplot2` to allow users to visualize multiple datasets in a unified layout. It supports **flexible customization of color schemes, thresholds, annotations**, and groupings to highlight molecular features (e.g., transcripts or proteins) that meet statistical criteria such as p-values and fold changes. One key feature is the ability to **explicitly show the number of features passing user-defined thresholds**, adding critical context often missing in conventional volcano or MA-plots.

This vignette showcases how to generate `ThresholdedScatterplot` visualizations from raw or summarized data and highlights its applicability to multi-condition, multi-group, or multi-timepoint designs. Compared to other Bioconductor scatter plotting tools, such as `EnhancedVolcano`, `ggVolcanoR`, or manual `ggplot2` solutions, `ThresholdedScatterplot` focuses on providing **quantitative clarity** through visual summary of significance thresholds and grouping structure.

It is particularly useful for **multi-condition** or **multi-group** designs, where the number of features passing user-defined thresholds can be explicitly shown. This is a critical feature that is often missing in conventional volcano or MA-plots.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Example Usage of ThresholdedScatterplot

This vignette demonstrates how to use `MultiModalGraphics` to create a `ThresholdedScatterplot` visualization.

```{r}
# Load required libraries
if (exists("suppressPackageStartupMessage", mode = "function")) {
  suppressPackageStartupMessage(library(MultiModalGraphics))
} else {
  library(MultiModalGraphics)
}

# Load the data
file_path <- system.file("extdata", "scatterexample.csv", package = "MultiModalGraphics")
filtered_data <- read.csv(file_path)

# Transform and set up for plot
filtered_data$negLog10p <- -log10(filtered_data$p)
filtered_data$negLog10q <- -log10(filtered_data$q)

# Create the ThresholdedScatterplot instance
scattered_plot <- ThresholdedScatterplot(
  data = filtered_data,
  logFoldChange = "log2fc",
  timePointColumn = "timePoint"
)

# Generate the plot
scattered_plot <- createPlot(
  scattered_plot,
  color1 = "cornflowerblue",
  color2 = "grey",
  color3 = "indianred",
  highLog2fc = 0.585,
  lowLog2fc = -0.585,
  expressionDirection = "regulation",
  timeVariable = "reg_time_org"
)

# Display the plot
scattered_plot
```
