---
title: "ThresholdedScatterplot MAE Example"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ThresholdedScatterplot MAE Example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Abstract

`ThresholdedScatterplot` is a highly customizable scatter plot class that extends `ggplot2` to allow users to visualize multiple datasets in a unified layout. It supports **flexible customization of color schemes, thresholds, annotations**, and groupings to highlight molecular features (e.g., transcripts or proteins) that meet statistical criteria such as p-values and fold changes. One key feature is the ability to **explicitly show the number of features passing user-defined thresholds**, adding critical context often missing in conventional volcano or MA-plots.

This vignette showcases how to generate `ThresholdedScatterplot` visualizations from raw or summarized data and highlights its applicability to multi-condition, multi-group, or multi-timepoint designs. Compared to other Bioconductor scatter plotting tools, such as `EnhancedVolcano`, `ggVolcanoR`, or manual `ggplot2` solutions, `ThresholdedScatterplot` focuses on providing **quantitative clarity** through visual summary of significance thresholds and grouping structure.

It is particularly useful for **multi-condition** or **multi-group** designs, where the number of features passing user-defined thresholds can be explicitly shown. This is a critical feature that is often missing in conventional volcano or MA-plots.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Example Usage of ThresholdedScatterplot with MAE

This vignette demonstrates how to use `MultiModalGraphics` to create a `ThresholdedScatterplot` visualization.

```{r}
# Load required libraries
if (exists("suppressPackageStartupMessage", mode = "function")) {
  suppressPackageStartupMessage(library(MultiAssayExperiment))
  suppressPackageStartupMessage(library(MultiModalGraphics))
  suppressPackageStartupMessage(library(dplyr))
} else {
  library(MultiAssayExperiment)
  library(MultiModalGraphics)
  library(dplyr)
}
# Parameters
num_genes <- 1000
num_samples <- 1000  # Total number of samples
conditions <- list(
  Condition1 = c("Cond1A", "Cond1B"),
  Condition2 = c("Cond2A", "Cond2B"),
  Condition3 = c("Cond3A", "Cond3B")
)

# Function to generate expression data
generate_expression_data <- function(num_genes, num_samples) {
  matrix(rnorm(num_genes * num_samples, mean = 6, sd = 2), nrow = num_genes, ncol = num_samples)
}

# Initialize lists
expression_data_list <- list()
metadata_list <- list()

# Generate samples for each combination of conditions
all_combinations <- expand.grid(conditions)
num_combinations <- nrow(all_combinations)
samples_per_combination <- num_samples %/% num_combinations
remaining_samples <- num_samples %% num_combinations

for (i in 1:num_combinations) {
  combination <- all_combinations[i, ]
  combination_name <- paste(combination, collapse = "_")

  # Calculate the number of samples for this combination
  num_samples_for_this_combination <- samples_per_combination + ifelse(i <= remaining_samples, 1, 0)

  expression_data <- generate_expression_data(num_genes, num_samples_for_this_combination)
  colnames(expression_data) <- paste0(combination_name, "_Sample", 1:num_samples_for_this_combination)
  rownames(expression_data) <- paste0("Gene", 1:num_genes)

  sample_metadata <- data.frame(
    SampleID = colnames(expression_data),
    Group = rep(c("Control", "Treatment"), length.out = num_samples_for_this_combination),
    stringsAsFactors = FALSE,
    TimePoint = rep(c("T1","T1" ,"T2",  "T2"), length.out = num_samples_for_this_combination),  # Example time points
    SampleType = rep(c("Tissue", "Organ"), length.out = num_samples_for_this_combination)  # Example sample types
  )

  for (cond in names(combination)) {
    sample_metadata[[cond]] <- combination[[cond]]
  }

  expression_data_list[[combination_name]] <- expression_data
  metadata_list[[combination_name]] <- sample_metadata
}

# Combine metadata into a single data frame
combined_metadata <- do.call(rbind, metadata_list)
rownames(combined_metadata) <- combined_metadata$SampleID

# Select only the relevant columns for the final metadata
final_metadata <- combined_metadata[, c("SampleID", "TimePoint", "SampleType", "Group", "Condition1", "Condition2", "Condition3")]

# Check the size of the final metadata
print(dim(final_metadata))  # Should be 1000 rows
head(final_metadata)

# Create the MultiAssayExperiment object
mae <- MultiAssayExperiment::MultiAssayExperiment(experiments = expression_data_list, colData = final_metadata)

# Check available assay names
print(names(MultiAssayExperiment::experiments(mae)))

# Use a valid assay name (example: select the first one)
assayNames <- names(MultiAssayExperiment::experiments(mae))
assayName <- assayNames[1]
print(paste("Using assay name:", assayName))

# Assuming ThresholdedScatterplot_MAE is already defined in your script
scatterplotObject <- ThresholdedScatterplot_MAE(
  mae = mae,
  assayName = assayName,
  timepoint = "TimePoint",
  sampleType = "SampleType",
  groupColumn = "Group"
)


# Assuming createPlot is already defined in your script
# Generate and display the plot
scatterplotObject <- createPlot(
  scatterplotObject,
  color1 = "cornflowerblue",
  color2 = "grey",
  color3 = "indianred",
  highLog2fc = 0.585,
  lowLog2fc = -0.585,
  negLog10pValue = 1.301
)
```

# Show the plot
```{r plot, fig.width=5, fig.height=4}
print(scatterplotObject@plot)

```
