
---
title: "Function Documentation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Function Documentation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

This vignette demonstrates how to use the `get_informative_heatmap_df()`, 
`get_clear_scatterplot_df()`, and `get_multifeature_grid_df()` functions from 
the MultiModalGraphics package.

## Informative Heatmap Data

The `get_informative_heatmap_df()` function provides a sample dataframe suitable 
for creating informative heatmaps. The dataframe includes values for three 
groups and their associated p-values, with row names indicating gene 
identifiers.

```{r informative-heatmap-example}
library(MultiModalGraphics)
informative_heatmap_df <- get_informative_heatmap_df()
print(informative_heatmap_df)
```

## Clear Scatterplot Data

The `get_clear_scatterplot_df()` function provides a sample dataframe that 
represents data suitable for creating a clear scatterplot. This includes log2 
fold change values, p-values, and regulation direction for a set of genes 
across different tissues and time points.

```{r clear-scatterplot-example}
clear_scatterplot_df <- get_clear_scatterplot_df()
print(clear_scatterplot_df)
```

## Multifeature Grid Data

The `get_multifeature_grid_df()` function returns a predefined dataframe that 
is ideal for creating multifeature grid visualizations. This dataframe includes 
signaling pathways, activation z-scores, and associated p-values for different 
tissues and time points.

```{r multifeature-grid-example}
multifeature_grid_df <- get_multifeature_grid_df()
print(multifeature_grid_df)
```

## Session info
```{r, echo=FALSE}
sessionInfo()
