
---
title: "Informative Heatmap Usage"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Informative Heatmap Usage}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ComplexHeatmap)
library(MultiModalGraphics)
```

# Informative Heatmap

This vignette demonstrates how to create and visualize an informative heatmap 
using the `MultiModalGraphics` package which utilizes the `ComplexHeatmap` 
package for enhanced visualizations.

## Basic Usage

Below is an example of how to create an informative heatmap with data 
representing genes, their value groups, and significance levels.

### Data Preparation

First, we load the necessary data. 

```{r load-data}
informative_heatmap <- get_informative_heatmap_df()
informative_heatmap_matrix <- as.matrix(informative_heatmap)
group_val <- informative_heatmap_matrix[, 1:3]
p_val <- informative_heatmap_matrix[, 4:6]
```

### Creating Heatmap

Create an `InformativeHeatmap` object with custom settings for visual 
representation.

```{r create-heatmap}
htmp_plus <- InformativeHeatmap(group_val,
                                unit_val = 7,
                                pch_val = 16,
                                significant_color = "black",
                                trending_color = "turquoise",
                                significant_pvalue = 0.05,
                                trending_pvalue = 0.1,
                                significance_level = p_val,
                                row_title = "Genes",
                                column_title = "Value and Significance",
                                cluster_rows = TRUE,
                                show_row_names = TRUE,
                                row_names_side = "left",
                                column_names_rot = 45,
                                row_dend_reorder = TRUE,
                                rect_gp = gpar(col = "white", lwd = 2))
```

### Drawing Heatmap

Finally, draw the heatmap using the ComplexHeatmap function to visualize the 
data.

```{r draw-heatmap}
draw(getHeatmapObject(htmp_plus))
```

## Conclusion

This vignette provided a simple example of how to use the `MultiModalGraphics` 
package to create a detailed heatmap for data analysis. For more advanced 
features and customization, refer to the `ComplexHeatmap` documentation.

## Session info
```{r, echo=FALSE}
sessionInfo()
