
---
title: "MultiModalGraphics: Comprehensive Guide"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MultiModalGraphics: Comprehensive Guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(MultiModalGraphics)
library(ComplexHeatmap)
library(dplyr)
```

# Introduction

The `MultiModalGraphics` package provides rich functionalities for combining different data types into an integrative visual representation. This package facilitates the creation of intuitive visualizations that merge multiple streams of data into a single coherent graphic.

This comprehensive guide includes the following sections:
- Generating Data
- Using ClearScatterplot
- Informative Heatmap Usage
- MultifeatureGrid Visualization

# Generating Data

This section demonstrates how to use the `get_informative_heatmap_df()`, `get_clear_scatterplot_df()`, and `get_multifeature_grid_df()` functions from the MultiModalGraphics package.

## Informative Heatmap Data

The `get_informative_heatmap_df()` function provides a sample dataframe suitable for creating informative heatmaps. The dataframe includes values for three groups and their associated p-values, with row names indicating gene identifiers.

```{r informative-heatmap-example}
informative_heatmap_df <- get_informative_heatmap_df()
print(informative_heatmap_df)
```

## Clear Scatterplot Data

The `get_clear_scatterplot_df()` function provides a sample dataframe that represents data suitable for creating a clear scatterplot. This includes log2 fold change values, p-values, and regulation direction for a set of genes across different tissues and time points.

```{r clear-scatterplot-example}
clear_scatterplot_df <- get_clear_scatterplot_df()
print(clear_scatterplot_df)
```

## Multifeature Grid Data

The `get_multifeature_grid_df()` function returns a predefined dataframe that is ideal for creating multifeature grid visualizations. This dataframe includes signaling pathways, activation z-scores, and associated p-values for different tissues and time points.

```{r multifeature-grid-example}
multifeature_grid_df <- get_multifeature_grid_df()
print(multifeature_grid_df)
```

# Using ClearScatterplot

This section demonstrates how to use the `ClearScatterplot` class to create enhanced scatterplots. The `ClearScatterplot` class provides functionalities for creating scatterplots with enhanced visualization features like coloring based on significance levels and plotting against different variables.

## Example: Creating a Scatterplot

First, load the necessary data and create an instance of `ClearScatterplot`.

```{r load-data}
plotdata <- get_clear_scatterplot_df()

scatterplotObject <- ClearScatterplot(
  data = plotdata,
  logFoldChange = "log2fc",
  timePointColumn = "timePoint",
  timePointLevels = c("T10R1", "T5R1")
)
```

Next, create the scatterplot using the `createPlot` method, specifying various parameters such as color and thresholds.

```{r create-plot}
scattered_plot <-
  createPlot(
    scatterplotObject,
    color1 = "cornflowerblue",
    color2 = "grey",
    color3 = "indianred",
    highLog2fc = 0.585,
    lowLog2fc = -0.585,
    negLog10pValue = 1.301,
    expressionDirection = "regulation",
    negativeLogPValue = "negLog10p",
    timeVariable = "reg_time_org",
    xAxis = "organ",
    yAxis = "timePoint"
  )
```

Finally, display the plot. This will call the 'show' method to render the plot.

```{r display-plot}
scattered_plot  # Display the plot
```

# Informative Heatmap Usage

This section demonstrates how to create and visualize an informative heatmap using the `MultiModalGraphics` package which utilizes the `ComplexHeatmap` package for enhanced visualizations.

## Basic Usage

Below is an example of how to create an informative heatmap with data representing genes, their value groups, and significance levels.

### Data Preparation

First, we load the necessary data.

```{r load-heatmap-data}
informative_heatmap <- get_informative_heatmap_df()
informative_heatmap_matrix <- as.matrix(informative_heatmap)
group_val <- informative_heatmap_matrix[, 1:3]
p_val <- informative_heatmap_matrix[, 4:6]
```

### Creating Heatmap

Create an `InformativeHeatmap` object with custom settings for visual representation.

```{r create-heatmap}
htmp_plus <- InformativeHeatmap(group_val,
                                unit_val = 7,
                                pch_val = 16,
                                significant_color = "black",
                                trending_color = "turquoise",
                                significant_pvalue = 0.05,
                                trending_pvalue = 0.1,
                                significance_level = p_val,
                                row_title = "Genes",
                                column_title = "Value and Significance",
                                cluster_rows = TRUE,
                                show_row_names = TRUE,
                                row_names_side = "left",
                                column_names_rot = 45,
                                row_dend_reorder = TRUE,
                                rect_gp = gpar(col = "white", lwd = 2))
```

### Drawing Heatmap

Finally, draw the heatmap using the ComplexHeatmap function to visualize the data.

```{r draw-heatmap}
draw(getHeatmapObject(htmp_plus))
```

# MultifeatureGrid Visualization

This section demonstrates how to use the `MultifeatureGrid` class from the `MultiModalGraphics` package to create comprehensive heatmaps integrating various data features such as z-scores, p-values, and counts.

## Example Data Preparation

First, we create a sample dataset representing biological data across different tissues and signaling pathways, with associated p-values and activation z-scores.

```{r example-data}
data <- get_multifeature_grid_df()
```

## Creating a MultifeatureGrid Object

We initialize the `MultifeatureGrid` object with the data prepared above.

```{r create-grid}
mg <- MultifeatureGrid(data)
```

## Plotting the Heatmap

We then plot the heatmap, specifying 'tissue' as the independent variable for faceting.

```{r plot-heatmap, fig.width=10, fig.height=8}
plot_heatmap(mg, independantVariable = "tissue")
```

This plot provides a visual summary of the signaling activity and the statistical significance across different tissues, utilizing a color gradient to represent activation z-scores and the size of points to indicate the number of genes involved.

# Conclusion

This comprehensive guide provided examples of how to generate data, create scatterplots, visualize informative heatmaps, and use multifeature grids using the `MultiModalGraphics` package. For more advanced features and customization, refer to the package documentation and the `ComplexHeatmap` and `ggplot2` documentation.

## Session info
```{r, echo=FALSE}
sessionInfo()
```
