
---
title: "MultifeatureGrid Visualization"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MultifeatureGrid Visualization}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(MultiModalGraphics)
```

# MultifeatureGrid Visualization

This vignette demonstrates how to use the `MultifeatureGrid` class from the `MultiModalGraphics` package to create comprehensive heatmaps integrating various data features such as z-scores, p-values, and counts.

## Example Data Preparation

First, we create a sample dataset representing biological data across different tissues and signaling pathways, with associated p-values and activation z-scores.

```{r example-data}
data <- get_multifeature_grid_df()

```

## Creating a MultifeatureGrid Object

We initialize the `MultifeatureGrid` object with the data prepared above.

```{r create-grid}
mg <- MultifeatureGrid(data)
```

## Plotting the Heatmap

We then plot the heatmap, specifying 'tissue' as the independent variable for faceting.

```{r plot-heatmap}
plot_heatmap(mg, independantVariable = "tissue")
```

This plot provides a visual summary of the signaling activity and the statistical significance across different tissues, utilizing a color gradient to represent activation z-scores and the size of points to indicate the number of genes involved.
